A Switching Planner for Combined Task and Observation Planning

### 2010-11-21 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/moritz_2011.tex
I have to go to a lunch. Writers block was just starting to subside too, what a shame...
 - Charles

RICHARD, I HAVE STOLEN A FEW LINES FROM YOUR HIPPO THREAD TO SKETCH
THIS STUFF. I HOPE IT'S TAKEN AS FLATTERY :)

We now describe a continual planning approach that switches between
sequential and decision-theoretic planning. The problem at hand is a
POMDP described in DTPDDL.  Our approach uses sequential planning to
compute a linear plan, and one trace of that plan's execution. The
latter encapsulates assumptions about: (1) the true underlying state,
(2) how the plan's execution will progress, and (3) the possibility of
eventually holding strong beliefs about the truth values of specific
state propositions. Sequential planning proceeds in a continual manner
until the validation of an assumption about a runtime state
proposition is scheduled. At that point the system switches to a
decision-theoretic planner that is used to tailor sensory processing
by the robot to test the validity of the assumptions so far.

The problem posed to the decision-theoretic system is an abstraction
of the problem at hand. It is induced according to the executing
sequential plan. Specifically, the abstract process is derived by
first omitting all problem variables that are not mentioned in the
linear trace, and then refining that preliminary model, by
progressively adding more problem variables while the resulting state
space remains sufficiently small. Refinement is a simple and
inexpensive step, guided by the level at which variable symbols occur
in the description of the initial conditions. The abstract process is
also given a utility model that captures the reward (resp. penalty) of
correctly (resp. falsely) in/validating the assumptions made in the
linear trace. In detail, for each assumption on a state variable, we
can have a reward of $1$ for correctly rejecting it, and a penalty of
$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
a more-or-less continual manner until an assumption is rejected, or
until it returns control to the sequential system because it can see
no benefit from performing further actions in the abstraction.

@@

\subsection{Sequential Planning in POMDP Determinisations}

\subsection{Abstractions for Decision-Theoretic Planner}

>>>>

We now describe a continual planning approach that switches between
sequential and decision-theoretic planning. The problem at hand is a
POMDP described in DTPDDL.

\subsection{Sequential Planning in POMDPs}

Our approach uses sequential planning to compute a linear plan, and
one trace of that plan's execution. The latter encapsulates
assumptions about: (1) the true underlying state, (2) how the plan's
execution will progress, and (3) the possibility of eventually holding
strong beliefs about the truth values of specific state
propositions. Sequential planning proceeds in a continual manner until
the verification of an assumption about a runtime state proposition is
scheduled. At that point the system switches to a decision-theoretic
planner that is used to tailor sensory processing by the robot to test
the validity of the assumptions so far.

The first actions sequenced by the planner are virtual, and correspond
to identifying an abstract start state. That prefix corresponds to a
{\em relaxed visitation} of the root term of the DTPDDL $(:init)$
block. In the relaxed case, we can modify the initial definition so
that a conjunctive term is visited provided all its atomic subterms
are visited, and zero or more of its probabilistic terms are
visited. Writing \#\ if the value of a proposition is unspecified by a
the {\em relaxed visitation}, the $\pp{R2D2}$ example has the
following abstract starting states:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
\hline
.24 & kitchen & office & office \\
.3 & kitchen & \# & office \\
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
.7 & kitchen & \# &  kitchen\\
.06 & kitchen & kitchen & office \\
.56 & kitchen & office & kitchen \\
.14 & kitchen & kitchen & kitchen \\
\hline
\end{tabular}
\normalsize

Sequential planning in our approach proceeds in two phases. In the
first phase, the goal is to identify a plan prefix

In sequential planning, the plan prefix corresponds to choosing a
valid abstract starting state.

moreover, for the purpose of decision-theoretic abstractions,

A relaxed form of {\em visitation} we call {\em abstract
  visitation}. In the abstract case a conjunctive term is visited
  provided all its atomic subterms are visited, and zero or more of
  its probabilistic terms are visited.

The planner has a compulsory starting action, that visits the root
term, and then has to execute visitation actions until that plan
prefix corresponds to a valid {\em abstract visitation} of the {\em
:init} block. Beyond such a prefix, the process actions and senses are
available to the planner

\subsection{Decision-Theoretic Planning in Abstractions}

The problem posed to the decision-theoretic system is an abstraction
of the problem at hand with a distinct utility model that focuses the
agents activities on assumption verification. The abstraction is
induced according to the executing sequential plan. Specifically, we
first omit all problem variables that are not mentioned in the linear
trace, and then refine that preliminary model, by progressively adding
more problem variables while the resulting state space remains of a
practicable size. Refinement in our approach is a simple and
inexpensive step, guided by the level at which variable symbols occur
in the description of the initial conditions. The abstract process is
also given a utility model that captures the reward (resp. penalty) of
correctly (resp. falsely) in/validating the assumptions made in the
linear trace. In detail, for each assumption on a state variable, we
can have a reward of $1$ for correctly rejecting it, and a penalty of
$-1$ for falsely rejecting it. Decision theoretic planning proceeds in
a more-or-less continual manner until an assumption is rejected, or
until it returns control to the sequential system because it can see
no benefit from performing further actions in the abstraction.

@@
